### 2017-03-07 manuscript/manuscript.tex
whoops. forgot these two things
all figures in finished form and added to manuscript
fixed error in temp-prec-disch plot, other plot mods, halfway through results

\usepackage{color, listings, bm, amsmath, blkarray}
% \usepackage[percent]{overpic}
\usepackage{tikz}
% \usepackage{graphicx}
% \usepackage{epsfig}

\begin{document}
% todo:
% include snowmelt in model run. see if it's worth including with precip and temp as a climate predictor
% introduce terminology:
% responses = water temp, Q
% climatic predictors = air temp, precip, drought
% landscape predictors = many
%include hydrograph figures from kathy liederman and julian olden to show
    %rain dom, rain and snow, snow dom
%include the over-time plots in the supplemental information
%in regression figure, line up left side of legend box with plot above
    %make gray horiz line go only to topright corner of box
    %make axis lines and plot borders cleaner (right now axis segments are darker)
    %instead of elevation on x, use ice and color by elev (orange to blue) X no story
    %for loadings, color by ice and make it darkgreen , midgreen, white X not without the above
    % add % ice as one of the hidden trends

@@

Potential climatic predictors of water temperature and discharge included mean and max air temperature, total precipitation, snowmelt, and hydrologic drought, averaged by month across the response variable time series. All but snowmelt are available through the U.S. Climate Divisional Dataset, developed by the National Centers for Environmental Information (NCEI)\colorbox{red}{\lstinline{cite}}. We acquired climatic predictor data grouped by Washington State climate division, and all but two of our sites fell within divisions 3 (Puget Sound Lowland) and 4 (East Olympic/Cascade Foothills). We therefore aggregated these data by monthly mean across the two regions (after verifying their post-standardization similarity), resulting in a single dataset of four regional, climatic predictors. A snowmelt time series was then added to this dataset, using monthly mean records from six SNOTEL sites (Bumping Ridge, Elbow Lake, Mount Crag, Park Creek Ridge, Stevens Pass, White Pass) listed by the USDA's Natural Resources Conservation Service \colorbox{red}{\lstinline{cite}}. We calculated monthly snowmelt for each site as the absolute value of negative differences in cumulative snow water equivalent from each month to the next. The snowmelt time series was assigned zeros for any positive differences (accumulations).

@@

Additional, non seasonal variation due to exogenous effects loads onto the shared trends, and a portion of remaining variation is absorbed by error matrix \textbf{v}. We fit models using four unique error structures (\textbf{R}), to allow for different suites of unknown drivers affecting rivers. We included shared variance and zero covariance, individual variance and zero covariance, shared variance and shared covariance, and unconstrained error. Details on these structures and their implications can be found in colorbox{red}{\lstinline{Holmes et al. (2012)}}. The best models for water temperature and discharge were determined with AIC.

@@

@@

% \begin{overpic}[grid,scale=0.5,unit=1mm]{figures/16_temp_all_reg.pdf}
% \begin{overpic}[width=1\textwidth,grid,tics=10]{figures/16_temp_all_reg.pdf}
% \put (1,55) {\small
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili\\
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili\\
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili\\
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili\\
% chili chili chili chili chili chili chili chili chili chili chili chili chili chili
% }
% \end{overpic}

% \begin{tikzpicture}
% \node (img) {\includegraphics[width=1\textwidth]{figures/16_temp_all_reg.pdf}};
% \node [below left,text width=4cm,align=left] at (img.north east){\small Lorem ipsum dolor sit amet consectetur and then a bunch more stuff that no one remembers.};
% \end{tikzpicture}

% \begin{picture}(100,100)
% \put(.1,.1){\includegraphics{figures/16_temp_all_reg.pdf}}
% % \put(.01\textwidth,.3\textwidth){hello}
% \end{picture}

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/16_temp_all_reg.pdf}
%   \caption{A boat.}
%   \label{fig:boat1}
% \end{figure}

% \begin{figure}[h!]
%    \centering
%    % \vspace{20pt}%
%    \includegraphics[width=\textwidth]{figures/16_temp_all_reg.pdf}
%    % \includegraphics[width=20pt]{figures/16_temp_all_reg.pdf}
%    % \caption{Epipolar Geometry}
%    % \label{fig:cclogo}
% \end{figure}

% \begin{figure}[h!]
% \begin{center}
% % \includegraphics[width=\textwidth]{figures/16_temp_all_reg.pdf}
% \resizebox{\textwidth}{!}{\includegraphics{figures/16_temp_all_reg.pdf}}
% \caption{Epipolar Geometry}
% \label{fig:cclogo}
% \end{center}
% \end{figure}

\begin{tikzpicture}
    \draw (0, 0) node[inner sep=0] {\includegraphics[width=\textwidth]{figures/16_temp_all_reg.pdf}};
    \draw (-4.1, 0) node[text width=10em] {\footnotesize \textbf{Figure 1} };
\end{tikzpicture}

@@

% https://wa.water.usgs.gov/data/realtime/adr/interactive/ #discharge daily
% https://waecy.maps.arcgis.com/apps/Viewer/index.html?appid=832e254169e640fba6e117780e137e7b #Q 15min
% http://www.horizon-systems.com/NHDPlus/NHDPlusV2_17.php #nhdplus
% ftp://newftp.epa.gov/EPADataCommons/ORD/NHDPlusLandscapeAttributes/StreamCat/HydroRegions/ #streamca
% ncdc.noaa.gov/cag/time-series/us #climate data
% https://www.nrcs.usda.gov/wps/portal/nrcs/detail/or/snow/?cid=nrcs142p2_046350 #snow
% https://arxiv.org/pdf/1509.00660.pdf #TMB
% Harvey, A. C. 1989. Forecasting, structural time series models and the Kalman flter. Cambridge University Press, Cambridge, UK.

>>>>

\usepackage{color, listings, bm, amsmath, blkarray, gensymb}
% \usepackage[percent]{overpic}
\usepackage{tikz}
% \usepackage{graphicx}
% \usepackage{epsfig}
% \usepackage{subfig}
% \captionsetup{belowskip=12pt}

\begin{document}
% todo:
% include snowmelt in model run. see if it's worth including with precip and temp as a climate predictor
% introduce terminology:
% responses = water temp, Q
% climatic predictors = air temp, precip, drought
% landscape predictors = many
%include hydrograph figures from kathy liederman and julian olden to show
    %rain dom, rain and snow, snow dom
%include the over-time plots in the supplemental information
%in regression figure, line up left side of legend box with plot above
    %make gray horiz line go only to topright corner of box
    %make axis lines and plot borders cleaner (right now axis segments are darker)
    %instead of elevation on x, use ice and color by elev (orange to blue) X no story
    %for loadings, color by ice and make it darkgreen , midgreen, white X not without the above
    % add % ice as one of the hidden trends 
%discharge regression combo plot is scrapped. no relationship with snowmelt?
%discharge only really relevant in the Q v temp plot.
%increase unit text size in fig 2
%change area over 1000m to % area over 1000m and replot
%change "dominated" to "dominant"
%change "fall" to "autumn"    
%dont size points in figure 2. include size in legend in fig 3

@@

Potential climatic predictors of water temperature and discharge included mean and max air temperature (\degree C), total precipitation (cm), snowmelt (cm), and hydrological drought (Palmer Hydrological Drought Index), averaged by month across the response variable time series. All but snowmelt are available through the U.S. Climate Divisional Dataset, developed by the National Centers for Environmental Information (NCEI)\colorbox{red}{\lstinline{cite}}. We acquired climatic predictor data grouped by Washington State climate division, and all but two of our sites fell within divisions 3 (Puget Sound Lowland) and 4 (East Olympic/Cascade Foothills). We therefore aggregated these data by monthly mean across the two regions (after verifying their post-standardization similarity), resulting in a single dataset of four regional, climatic predictors. A snowmelt time series was then added to this dataset, using monthly mean records from six SNOTEL sites (Bumping Ridge, Elbow Lake, Mount Crag, Park Creek Ridge, Stevens Pass, White Pass) listed by the USDA's Natural Resources Conservation Service \colorbox{red}{\lstinline{cite}}. We calculated monthly snowmelt for each site as the absolute value of negative differences in cumulative snow water equivalent from each month to the next. The snowmelt time series was assigned zeros for any positive differences (accumulations).

@@

Additional, non seasonal variation due to exogenous effects loads onto the shared trends, and a portion of remaining variation is absorbed by error matrix \textbf{v}. We fit models using four unique error structures (\textbf{R}), to allow for different suites of unknown drivers affecting rivers. We included shared variance and zero covariance, individual variance and zero covariance, shared variance and shared covariance, and unconstrained error. Details on these structures and their implications can be found in \colorbox{red}{\lstinline{Holmes et al. (2012)}}. The best models for water temperature and discharge were determined with AIC.

@@

For post-hoc analyses, monitoring sites were separated into three classes based on relative areal coverage of perennial ice and/snow (hereafter ``\% glaciation'') and mean elevation across their watersheds. The three classes are loosely based on the classification scheme and language of the Climate Impacts Group at the University of Washington \colorbox{red}{\lstinline{cite}}, and are here delineated according to Table 1.

% \begin{table}
\begin{center}
\textbf{Table 1} Watershed classification scheme
\end{center}
\begin{center}
\begin{tabular}{ |c|c|c|c| } 
 \hline
 Classification & Abb. & Glaciation (\%) & Mean elev. (m) \\
 \hline
 Rain-dominated & RD & $< 0.7$ & $< 600$ \\
 Rain-and-snow & RS & $< 0.7$ & $\geq 600$ \\
 Snow-dominated & SD & $\geq 0.7$ & - \\
 \hline
\end{tabular}
\end{center}
% \end{table}

@@

Mean monthly temperature trends for the three river classes, aggregated across all 38 years of data, deviated by a minimum of 1.0\degree C in December, and a maximum of 3.9\degree C in July (Fig. 1). SD rivers remained approximately two degrees colder than their RS counterparts through mid-late summer, and 3-4 degrees colder than RD throughout spring and summer. RD rivers were consistently warmest throughout the year. In January, RS reached a minimum of 4.4\degree C, and did not significantly differ from SD \colorbox{red}{\lstinline{stats throughout this section}}. RD, meanwhile, only attained a minimum of 5.6\degree C. RS reached a peak summer temperature of 16.9\degree C in July, while RS and SD followed in August with peak temperatures of 15.5 and 13.5\degree C, respectively.

Meanwhile, the amplitude of T\textsubscript{air} oscillation exceeded that of any river class, dipping below T\textsubscript{water} in autumn to a minimum of 3.2\degree C in December, and rising above RS and SD in March to an August maximum of 17.4\degree C. T\textsubscript{air} did not overtake RD T\textsubscript{water} until August, by which time the latter had begun to decline.

% \begin{figure}[h!]
\begin{center}
% \includegraphics[width=\textwidth]{figures/16_temp_all_reg.pdf}
\resizebox{\textwidth}{!}{\includegraphics{figures/12_air_water_discharge.pdf}}
% \caption{Epipolar Geometry}
% \label{fig:air_water_disch}
\end{center}
% \end{figure}
\textbf{Figure 1} Monthly mean T\textsubscript{water} by river class, and T\textsubscript{air} and Q across classes, from 1978 to 2015. All depicted series represent discrete data. \colorbox{red}{\lstinline{see appendix for standard deviations?}} \\[\baselineskip]
% The combined hydrograph of all rivers reveals two primary peaks, one beginning in late spring and the other extending from late fall to early winter, with a prominent trough in late summer. Spring peak discharge coincides noticeably with a separation in water temperature between SD and RS, while the summer trough coincides with separation of RD and T\textsubscript{air}. November marks both the autumn peak in discharge and the point at which T\textsubscript{air} falls below T\textsubscript{water}.

DFA results, aggregated across months and years for each site, revealed a trend toward reduced $T_{air}\rightarrow T_{water}$ coupling with greater watershed elevation ($p=0.04, R^2=0.18$; Fig. 2a). On average, a 1\degree C change in T\textsubscript{air} corresponded to a $0.53\pm 0.03$\degree C change in T\textsubscript{water} at RD, a $0.51\pm 0.08$\degree C change at RS, and a a $0.45\pm 0.17$\degree C change at SD sites. A similar trend was observed with respect to $Precip\rightarrow T_{water}$ coupling ($p=0.03, R^2=0.21$; Fig. 2b), where a monthly change in total precipitation of 1 cm corresponded to a $0.02\pm 0.009$\degree C change in T\textsubscript{water} for RD, $-0.003\pm 0.009$\degree C for RS, and $0.004\pm 0.02$\degree C for SD. There was no evidence of coupling between snowmelt and T\textsubscript{water} (Fig. 2c). It is of note that the strongest examples of $T_{air}\rightarrow T_{water}$ and $Precip\rightarrow T_{water}$ coupling were observed in the Duckabush River, while the weakest examples are from the Elwha River. Both rivers drain glaciers of the Olympic Mountain Range, and both are SD, though the Elwha's watershed is larger.

\begin{tikzpicture}
    \draw (0, 0) node[inner sep=0] {\includegraphics[width=\textwidth]{figures/16_temp_all_reg.pdf}};
    \draw (-4.1, -1.5) node[text width=9.5em] {\normalsize \textbf{Figure 2} Relationships between watershed elevation and climatic effects on T\textsubscript{water} \textbf{(a-c)}, and other watershed predictors and factor loadings on shared trends \textbf{(d-g)}. Regression lines indicate slopes significant at $\alpha=0.05$.};
\end{tikzpicture}

\begin{center}
\resizebox{\textwidth}{!}{\includegraphics{figures/15_disch_vs_air_focal5.pdf}}
\textbf{Figure 1} chili chili
\end{center}

Body text

\section*{Discussion}
The relationship between river classification and factor loadings is less clear

@@

% https://wa.water.usgs.gov/data/realtime/adr/interactive/ #discharge daily
% https://waecy.maps.arcgis.com/apps/Viewer/index.html?appid=832e254169e640fba6e117780e137e7b #Q 15min
% http://www.horizon-systems.com/NHDPlus/NHDPlusV2_17.php #nhdplus
% ftp://newftp.epa.gov/EPADataCommons/ORD/NHDPlusLandscapeAttributes/StreamCat/HydroRegions/ #streamca
% ncdc.noaa.gov/cag/time-series/us #climate data
% https://www.nrcs.usda.gov/wps/portal/nrcs/detail/or/snow/?cid=nrcs142p2_046350 #snow
% https://arxiv.org/pdf/1509.00660.pdf #TMB
% Harvey, A. C. 1989. Forecasting, structural time series models and the Kalman flter. Cambridge University Press, Cambridge, UK.
%CIG puget sound state of knowledge http://cses.washington.edu/picea/mauger/ps-sok/ps-sok_sec03_watercycle_2015.pdf
